---
title: "test"
author: "Efrat Vilenski"
date: "December 11, 2015"
output: slidy_presentation
---

## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Code and Output

```{r}
summary(cars)
```

## Slide with Plot

```{r, echo=FALSE}
plot(cars)
```


## Random
```{r, fig.width=8, fig.height=4}
library(parcoords)
set.seed(10)
a <- rnorm(50,0,1)
b <- rnorm(50,0,1)
c <- rnorm(50,0,1)
d <- rnorm(50,0,1)
e <- rnorm(50,0,1)
data <- data.frame(a=a, b=b, c=c, d=d, e=e)    
parcoords(data, rownames=FALSE,reorderable = T,brushMode = "2d-strums")

```

## Random2
```{r, echo=FALSE, results='asis', warning=FALSE}
plot(cars)

```

## Random
```{r, fig.width=8, fig.height=4}
set.seed(10)
a <- rnorm(50,0,1)
b <- rnorm(50,0,1)
c <- rnorm(50,0,1)
d <- rnorm(50,0,1)
e <- rnorm(50,0,1)
data <- data.frame(a=a, b=b, c=c, d=d, e=e)    
parcoords(data, rownames=FALSE,reorderable = T,brushMode = "2d-strums")

```

