# Pre-Processing

MultiNav contains default options for anomaly scores for the novice analyst, but also enables the advanced analyst to provide custom calculated scores. The package also includes several functions for common data pre-processing tasks (used internally by MultiNav package) that can be used independently from the visualizations. 


## Univariate Matrix

**Input:** 
The function expects an input dataset in the format of data.table, data.frame or matrix which contains continuous variables from the same scale.
A unique numeric row name should be assigned to each observation and will be used as unique identifier (id). If row names were not assigned, row number will be used as id.

A unique **numeric** column name should be assigned to each variable.
Any missing data should be handled before calling the MultiNav() function. Note: Any records containing missing values will be omitted by default.


**Output:**
Dataset with collection of univariate descriptive statistics performed on the values of each **variable**.


Statistic  | Description
------------- | -------------------------------------------
min  | Minimum value.
quantile_25  | Quantile 25.
median  | Median value.
quantile_75  | Quantile 75.
max  | Maximum value.
mean  | Mean value.
sd  | Standard deviation.
mad  | Mad - Median Absolute Deviation from the median. Robust estimate of standard deviation.
Skewness  | A measure of the asymmetry.
Kurtosis  | A measure of the "tailedness".

<br/>

**Example** 
```{r, message=FALSE}
library(MultiNav)
data <- DendrometerSensors
```

```{r}
uni_matrix<-Calc_uni_matrix(data)
head(uni_matrix)

```


## Quantiles Matrix
**Input:** 
The function expects an input dataset in the format of data.table, data.frame or matrix which contains continuous variables from the same scale.
A unique numeric row name should be assigned to each observation as unique identifier. If row names were not assigned, row number will be used as id.
Any missing data should be handled before calling the function. Note: Any records containing missing values will be omitted by default.


**Output:**
Dataset with collection of five selected quantiles performed on each row (observation) of the data (5, 25, 50, 75 and 95 are the default quantiles). minimum and maximum of each row are also included. The 'quantiles matrix' is set in the exact format needed as input for the functional boxplot chart.   

First Header  | Second Header
------------- | -------------
min  | min value
quantile_a  | by default = Quantile 05.
quantile_b  | by default = Quantile 25.
median  | Median value.
quantile_c  | by default = Quantile 75.
quantile_d  | by default = Quantile 95.
max  | max value


**Example - Defalt Quantiles** 

```{r q1}
quantiles_matrix<-Calc_quantiles_matrix(data)
head(quantiles_matrix)

```

**Example - Custom Quantiles** 


```{r q2}
quantiles_matrix<-Calc_quantiles_matrix(data, quantiles = c(0.02,0.2,0.8,0.98))
head(quantiles_matrix)

```


## Anoamly Scores
MultiNav provides several methods for calculating multivariate anomaly scores. The package provide only few methods, out of many other available methods (this is a popular area of research). We aimed to find methods that will likely be suitable for most use cases (that follow the data assumptions outlined). 
**Note** Selecting the right multivariate anomaly methods for each dataset needs to be done with care.  

### Hotelling's $T^2$ 
Hotelling's $T^2$ is a popular multivariate method for anomaly detection. Calculating $T^2$ requires estimation of the covariance matrix. MultiNav provides several variations of Hotelling's $T^2$ that are based on different methods of estimating the covariance matrix. More on Hotelling's $T^2$ and the selected variations in our upcoming paper... 

```{r fig.height=3.5, message=FALSE}
data<-t(LambsWeight[60:62,])
T2 <- round(T2(data),2)
T2_spearman<-round(T2_spearman(data),2)
T2_mcd50<-round(T2_mcd(data),2)
T2_mcd75<-round(T2_mcd75(data),2)
T2_mve<-round(T2_mve(data),2)
T2_CrouxOllerer<-round(T2_CrouxOllerer(data),2)
T2_SrivastavaDu<-round(T2_SrivastavaDu(data),2)

Anomaly.Scores<- cbind(sensorID=as.numeric(row.names(data)), 
                 T2,T2_spearman,T2_mcd50,T2_mcd75,T2_mve,
                 T2_CrouxOllerer,T2_SrivastavaDu)
head(Anomaly.Scores)

```


### Eigenvector Centrality 
Another multivariate anomaly scoring method is based on Eigenvector centrality, which is a popular network centrality measure, calculated based on pearson correlation estimate. This method can be efficient for finding anomalies in data with high correlation levels (such as in our first example of dendrometers sensors). 
```{r fig.height=3.5, message=FALSE}
data <- DendrometerSensors
EigenvectorCentrality<-EigenCentrality(data)
plot(EigenvectorCentrality,  xlab="Sensor ID",pch = 19)
```


