# Snapshot Scores View

The "snapshot scores" is MultiNav's main display for common anomaly detection tasks. In this section, we will review and demonstrate the various parameters of the display. In addition, more demonstrations are provided in the example sections.

In this display, anomaly scores are presented to the analyst for inspection, each anomaly score is displayed in a separate scatter plot. Each point in the scatter plot represents a specific sensor. When the analyst's hoovers over a specific sensor, the matching scores (of the same sensor) from all the other anomaly scores are highlighted (with a larger point and orange color). In addition, the values from the selected sensor are presented layed on top of a functional box plot, which enables to compare the sensor's readings relative to readings from the other sensors. This display relays on interactivity of linked charts to enable quickly examination of different sensors.

The base syntax:
<p class="well" style=" background-color: #F1F1F1; padding: 10px; border: 1px solid #cccccc;">MultiNav(data,type = "snapshot scores")</p>


Loading the package and some data.
```{r sv1, warning=FALSE, message=FALSE}
library(MultiNav)

data <- DendrometerSensors 

```

## Default Scores
MultiNav provides as default, a set of first order anomaly scores per variable (Quantile 25, Quantile 75, median and mad). This set of default scores is suitable for initial exploratory data analysis (EDA) and finding univariate anomalies.

**Input:** 
The only input required is a continuous multivariate dataset with variables on the same scale. The dataset should be in the format of a data.table, data.frame or matrix. Observations with NA values will be omitted. 


```{r sv2, fig.height=5.5}
MultiNav(data, type = "snapshot scores")
```

<br/><br/>

## Default Multivariate Scores 

MultiNav also provides two sets of multivariate default scores. We tried to find usefull defaults, but unlike the first order scores default, the usefullness of the methods depends on the input dataset charactaristics. 


### MV Default
Includes two selected methods of multivariate scoring: 

+ T2_MCD75 - A robust variation the popular Hotelling's T2 (known from multivariate process control literature), which is the multivariate version of the univariatet-test. The variation is based on the robust covariance estimation with "Minimum Covariance Determinant" (MCD) with breaking point of 75%.
+ EigenvectorCentrality - a popular network centrality method that can be applied on estimated distance matrix. The eigenvector centrality of each sensor is its summed connections to
other sensors, weighted by their centralities. This method is useful in dataset with high correlation levels between sensors (as exhibited in the first example section).

```{r sv6a}
data<-LambsWeight

window<-3
data.3D<-data[(dim(data)[2]-window+1):dim(data)[2],]

MultiNav(data.3D,type = "snapshot scores", 
         scores = "mv")
```
<br/><br/>

### T2 Variations
Another set of default scores, contains different variations of Hotelling's T2. T2 score is based on covariance matrix estimation, which can be challenging task if outliers are present or there is not enough data. There are many available methods for estimating the covariance matrix. Choosing the right one is crucial for achieving accurate T2 scores. This default set of scores provides several variations for calculating T2. Typically, analyst will examine the scores and decide which one of them is suitable for the dataset in question.

```{r sv6}
MultiNav(data.3D,type = "snapshot scores", 
         scores = "T2_Variations")
```
<br/><br/>


## Custom Scores
MultiNav also alows input of custom scores per sensor.

**Input:** Continuous multivariate dataset and a dataset of scores for each sensor (up to four different scores). 

```{r sv3, fig.height=5.5}

quantile_25 <- apply(data,2,quantile,probs=0.25,na.rm = TRUE)
median <- apply(data,2,median,na.rm = TRUE)
quantile_75 <- apply(data,2,quantile,probs=0.75,na.rm = TRUE)
mad<- apply(data,2,mad,na.rm = TRUE)
anomaly.scores<- cbind(id=as.numeric(colnames(data)), 
                 quantile_25,median,quantile_75,mad)
head(anomaly.scores)

MultiNav(data,type = "snapshot scores", scores = anomaly.scores)

```

<br/><br/>



## Custom Quantiles
MultiNav also alows input of custom quantails, used for the functional box plot background.

**Input:** Continuous multivariate dataset + Pre calculated quantiles dataset (in the structure detailed in the pre-processing section). The quantiles will be used for the functional box plot (instead of the default quantiles).

```{r sv5, fig.height=5.5}
quantiles_matrix<-Calc_quantiles_matrix(data)
head(quantiles_matrix)

MultiNav(data,type = "snapshot scores", quantiles=quantiles_matrix)

```
<br/><br/>



## Examine Historical Scores
MultiNav also enables to review historical scores per sensor (calculated per "Sliding Temporal Window").

**Input:** Continuous multivariate data. The window parameter is set, and historical scores per sliding window are calculated.

Note: since scores are calculated for each sensor per each sliding window, this action maybe time consuming! 

```{r sv7, fig.height=5.5}

MultiNav(data,type = "snapshot scores", 
         scores = "mv",window=3)
```
<br/><br/>


There is also an option for providing calculated scores per sensor per window. The calculated scores dataset should be in the specific format as in the example below.
When note - when providing calculated scores per temporal window, the 'window' parameter does not need to be set.


```{r sv7a, fig.height=5.5}

scores<-sliding_window_scores(data,14,scores="T2_Variations") #
head(scores)
MultiNav(data,type = "snapshot scores", sliding_scores = scores)

```
<br/><br/>



## Examine Outliers
**Input:** Continuous multivariate data. 

```{r sv8, fig.height=5.5}

MultiNav(data, type = "snapshot scores", 
         scores = "mv",window=3, show_diff = TRUE)

```
<br/><br/>
