---
output: html_document
editor_options: 
  chunk_output_type: console
---
# Example I - Dendrometer Sensors {#example1}

The first example corresponds to the task of identifying  malfunctioning [dendrometer sensors](https://en.wikipedia.org/wiki/Dendrometry) with unsupervised anomaly detection methods. Dendrometers are measurement devices that continuously monitor size variation in plant organs like stems, branches and fruits. Data from dendrometers can be used for monitoring plants' daily water status, and make automated irrigation decisions. A major challenge is to identify malfunctioning sensors and filter them out before making irrigation decisions.

</br>
Loading the necessary packages:
```{r, message=FALSE, warning=FALSE}
library(MultiNav)
library(RColorBrewer)
library(data.table)
```

## Exploratory Data Anaylsis with Univariate Scores

The dendrometer dataset includes data of 100 sensors from 120 hourly readings (5 days). 
We start by loading the 'DendrometerSensors' dataset and performing initial exploration to get a feel of how the data looks like and to check if there are any sensors that have obvious anomalies. 

The exploration starts with a simple heatmap. It provides a quick view of the data. 
Daily growth cycles are clear. 
Some anomalies can be seen, but it is hard to understand the details. 
[TODO: add some discussion of what you see.]

```{r ex1, message=FALSE, warning=FALSE}
data <- DendrometerSensors 

cols <- rev(colorRampPalette(c("red", "white", "blue"))(100)) # make color palette

lattice::levelplot(as.matrix(data), 
                   col.regions = cols[length(cols):1],
                   ylab="Dendrometer Id",xlab="Hours") # heatmap
```

To idetify anomalous sensors, and find more subtle anomalies than the ones visible in a heatmap, we will use  __MultiNav__. 
We start with a battery of univariate, sensor-wise, summary statistics.
The univariate summary statistics that are chosen by default are the 0.25,0.5,0.75 quantiles, and the MAD.

Hoovering over a sensor in the scatter plots will automatically highlight the sensor in all plots. It will also show the sensor's raw measurements on a _functional box-plot_.
With this display, we can visually identify several anomalous sensors (104, 142, 161, 182) and investigate the nature of the anomaly, with the functional boxplot.

```{r ex2, message=FALSE, warning=FALSE}
MultiNav(data, type = "snapshot scores")
```
[TODO: "snapshot scores" is a long name. Use one word. Maybe "snapshot", "snap", "main"? What other options are available except "snapshot"?]


</br></br>


## Evaluate Multivariate Scoring Methods
The previous univariate summaries, are computed based on the whole data. 
To study local patterns, we may want to compute summaries in a sliding window. 
Moreover, we may want to account for correlations in each window. 
The following scoring methods, collectively called "multivariate", first estimate the correlations within a single window, and then compute an anomaly score for the window. 
One may think of multivariate, as "second order" scores, because we will use an estimate of the correlation between periods.

We propose various methods for computing correlations, and scores. 
Readers familiar with the multivariate testing literature, will recognize that our various multivariate scores are all variations on Hotelling's classic $$T^2$$ statistic.
The appropriate method depends on the size of the window, the number of sensors, and the empirical correlations in the data.
For this reason, we do not recomend a-priori a single method, but rather, supply the user various such scoring rules. 

[TODO: use proper indentation for functions, and R object: `data.table`, `Calculated.Scores`, etc.]
By default, we use windows of 72 hours [TODO: the default window size needs to be a function of the data. Maybe sqrt of the number of periods? I also see that this is not actually a default, but rather, an arbitrary choice you made]. 
As pre-processing for this step, it is required to remove hours where the standard deviation equals 0 (which leaves 66 valid hours). [TODO: please automate this verificaton within your functions, instead of asking the user to do it]
The anomaly scores per sensor are calculated and stored in a `data.table` called `Calculated.Scores`, and passed to the `MultiNav()` function using the `scores=`  argument.

[TODO: some detail of the default multivaraite scoring rules is necessary.]

[TODO: use comments to exaplain what your code is doing]
```{r e36, message=FALSE, warning=FALSE, error=FALSE, fig.height=5.5}
mat<-data[49:120,] # subset last 72 hours 
exclude.hours<- c(1,24,25,48,49,72) #[TODO: make the extraction automated].
mat<-mat[-exclude.hours,]

set.seed(123)
T2 <- T2(t(mat)) # [TODO: avoid replacing an object in place.]
T2_mcd75 <- T2_mcd75(t(mat))
T2_co <- round(T2_CrouxOllerer(t(mat)),5)
EigenCent<-EigenCentrality(mat) # [TODO:loading a package from within a function is bad style.]

Calculated.Scores <- as.data.table(cbind(
  id=as.numeric(colnames(mat)), 
  T2=T2,
  T2_mcd75=T2_mcd75, 
  T2_co=T2_co, 
  EigenCent))
head(Calculated.Scores) # Inspect customize scores

MultiNav(mat, type = "snapshot scores", 
         scores = Calculated.Scores) # [TODO: why do I get a warning? #TODO: why am I getting a warning message?]

``` 
</br></br>


After visual examination of the anomaly scores, we can see:

+ Some sensors are identified as anomalous by viewing the multivariate scores, but they were not detected by the univariate scores. For example sensor 150.

+ _T2 score_ - This score does not seem to separate the anomalies as cleanly as the robust _T2_MCD75_. This is probably due to the fact that the covariance within a window is estimated from multiple windows: because some windows contain anomalies, then _masking_ occurs. A robust scoring rule such as _T2_MCD75_ is not masked, and is thus more sensitive to outliers. 

+ _EigenvectorCentrality_ score is good at capturing anti-correlation patterns, and works well in this use case as complimentary method to _T2_MCD75_. _EigenvectorCentrality_ identified sensor 104 as outlier and also identified sensor 187 which were not identified by _T2_MCD75_.  

+ _T2_co_ score had great potential: being designed to be both robust, and adaptive to high-dimension. Alas, in our specific example, it is a poor performer [TODO: explain why].


## Review Outliers
With univariate scoring rules, it is easy to understand the source of an anomaly. With windowed multivaraite rules, the nature of an anomaly is less obvious. 
We thus want to investigate anomalies detected. 
For this purpose we use the "snapshot scores" display again. 
This time we will provide as input, a longer data stream containing data from 120 periods, not only 73. 
In this period, we calculate historical scores in windows of size 66. 
We use `scores="mv"` to call for the default multivariate scores,: _T2_MCD75_ and _EigenvectorCentrality_.
[TODO: you should start with presentng scores="mv", and then show how to compute manually the scores matrix.]

```{r e4, message=FALSE, error=FALSE, warning=FALSE}
exclude.hours<- c(1,24,25,48,49,72,73,96,97,120) #[TODO: never extract manually. Always use a script to generate the extraction indices.]

MultiNav(data[-exclude.hours,], type = "snapshot scores", window=66, scores="mv")
``` 
[TODO: this display takes a while to load. Can you make it more efficient?]

</br></br>

Looking at the display, and hovering over a scatter plot point, we can learn more about the strength and weakness of our two selected methods for identifying specific patterns of anomalies. 

In general, we can visually observe that the _T2_mcd75_ separates the anomalous sensors better than _EigenvectorCentrality_, but it misses some anomalous sensors such as 104, 187. 

To review the historical scores of a specific sensor, click on the sensor's point in one of the scores scatter plots. The displays updates, the focus moves to the selected sensor and the historical scores for the sensor are shown. This functionality enables to review if there is a trend of improvement or deterioration in the sensor's historical scores. 

