# Example II - Lambs Weight

The second example dataset corresponds to a challenge of identifying anomalous lambs (that may require farmers attention) based on daily weight estimates of individual lambs in a herd. Our goal is to identify anomalous lambs and provide visual tools to examine and understand the anomalies. 

The dataset contains data of 80 daily weight estimates for 39 lambs. Data was collected from a novel system (developed by the Volcani PLF lab) which contains electronic scales and drinking behavior sensor, designed for automatic small-ruminant monitoring. The water source is an attraction point, that voluntarily attracts the animals to the scale several times per day.

</br>
Loading the necessary packages:
```{r, message=FALSE}
library(MultiNav)
library(corrplot)
library(RColorBrewer)


```

##  Initial Exploration

First we will load and examin the lambs dataset with a basic heat map. We can see that during the 80 days, the lambs gain weight rapidly.

```{r ex0a, fig.height=8, fig.width=10}
data<-LambsWeight

cols <- rev(colorRampPalette(brewer.pal(6, "Blues"))(100))
lattice::levelplot(t(data), col.regions = cols[length(cols):1], ylab="Days",xlab="Lamb ID")

```
</br></br>

Then we explore the data for the whole period of 80 days with MultiNav's default first order scores. Each point in the scatter plot, represents a lamb.

```{r ex0b}
MultiNav(data,type = "snapshot scores")

```
</br></br>

Next we will calculate and examine the daily weight change in percentage. The anomaly scores that will be calculated in the next sections will also be based on the weight gaps.

```{r ex4}
data_diff<-apply(data, 2, diff)
data_diff_per<-100*data_diff/data[1:dim(data)[1]-1,]
MultiNav(data_diff_per,type = "snapshot scores")

```
</br></br>

From both of the exploration displays, we can get see that during the whole period, there are no extream outliers lambs manifested in the first order scores.

## Examin Multivariate Outliers
In this use case, we demonstrate additional MultiNav functionality. The goal is to find individual outlier lambs and understand the nature of the anomaly. The anomalies will be examined in diffrent time windows.

### Three Days Window
Here we calculate an anomaly score for each lamb, based on variations of T2 Hotelling method. The parameter "show_diff = TRUE" is set, which adds an additional functional box plot of the differenced data. This plot will enable to investigate further the nature of the anomaly. For example, the anomaly in sensor 692 is easy to understand as it is manifested in each one of the days. But for sensor 628, the anomaly is not manifested in any specific day. To understand the nature of the anomaly, we will look at the added differenced data functional boxplot. We see that there a rapid change in the differenced data which is much more than the difference in the other normal sensors.

Additionally, the parameter "window=3" was added. When using the "window" parameter, we provide longer stream of data and scores per sliding window are calculated (beware, this may be computationally expensive). Only scores from the last window are initially displayed, but when clicking on a point representing a sensor, the historical scores for the sensor are displayed.
```{r ex4a, fig.height=5.5}

MultiNav(data_diff_per,type = "snapshot scores",
         scores = "T2_Variations",window=3, show_diff = TRUE)

``` 

</br></br>


### 7 Days Window
Next we will calculating anomaly score for each lamb, using the "mv" desult scores set. Two scores are calculated, T2_MCD75 which seems to work well, and the EigenvectorCentrality methods, which does not sepereate anomalies well for this dataset.

```{r ex4d, fig.height=5}

window<-7
data.last.7D<-data_diff[(dim(data_diff)[2]-window+1):dim(data_diff)[2],]


MultiNav(data_diff_per,type = "snapshot scores", scores = "mv",window=7, show_diff = TRUE)

``` 
</br></br>



