---
output: html_document
editor_options: 
  chunk_output_type: console
---
# Example II - Lambs Weight

The second example corresponds to identifying anomalous lambs, that may require farmers attention, based on daily weight estimates of individual lambs in a herd. Our goal is to identify anomalous lambs and provide visual tools to examine and understand the anomalies. 

The dataset contains data of 80 daily weight estimates for 39 lambs. Data was collected from a novel system (developed by the Volcani PLF lab [TODO: add link]) which contains electronic scales and drinking behavior sensor, designed for automatic small-ruminant monitoring. The water source is an attraction point, that voluntarily attracts the animals to the scale several times per day. [TODO: maybe add an image?]

</br>
Loading the necessary packages:
```{r, message=FALSE}
library(MultiNav)
library(corrplot)
library(RColorBrewer)
```

##  Initial Exploration

First we will load and examine the lambs dataset with a basic heat map. We can see that during the 80 days, the lambs gain weight rapidly.

```{r ex0a, fig.height=8, fig.width=10}
data<-LambsWeight

cols <- rev(colorRampPalette(brewer.pal(6, "Blues"))(100))
lattice::levelplot(t(data), 
                   col.regions = cols[length(cols):1], 
                   ylab="Days",
                   xlab="Lamb ID") # TODO: time usually on x axis. 
```
</br></br>

Then we explore the data for the whole period of 80 days with MultiNav's default univariate scores. Each point in the scatter plot, represents a lamb.

```{r ex0b}
MultiNav(data,type = "snapshot scores") # TODO: hover. not hoover. 
```
</br></br>

From the functional boxplot, we see a general grow trend. 
Because we prefer data that is statistically "stable", we will examine the growth-ratio. The anomaly scores that will be calculated in the next sections will also be based on the weight gaps.

```{r ex4}
data_diff<-apply(data, 2, diff)
data_diff_per<-100*data_diff/data[1:dim(data)[1]-1,]
MultiNav(data_diff_per,type = "snapshot scores")
```
</br></br>

From both of the exploration displays, we can get see that during the whole period, there are no extreme outliers lambs manifested in the univariate scores.

## Examin Multivariate Outliers
We now proceed from univariate summary statistics, to a multivariate score, in a short time-window. 

### Sliding Window of Univariate Scores
We now move from univariate scores in a window, to multivariate scores, in a window. As seen in Example @\ref(example1), our multivariate scores are variations of Hotelling's classical $$T^2$$ statistic.
The parameter `show_diff = TRUE` adds an additional functional box plot of the differenced data.The anomaly in sensor 692 is easy to understand as it is manifested in each one of the days. But for sensor 628, the anomaly is not manifested in any specific day. To understand the nature of the anomaly, we will look at the functional boxplot of differenced data. We see that there a rapid change in the differenced data which is much more than the difference in the other normal lambs.

Additionally, the parameter `window=3` was added. When using the `window` parameter, instead of a single summary for the whole data, we compute the summary in a sliding window (beware, this may be slow). Only scores from the last window are initially displayed, but when clicking on a point representing a sensor, the historical scores in the sliding window are displayed. [TODO: this is redundant. Should have been explained in the first example.]
One may return from the single sensor view, to the all-sensor view, by hitting the "Back to scores" button (upper left corner).
```{r ex4a, fig.height=5.5}
MultiNav(data_diff_per, type = "snapshot scores",
         scores = "T2_Variations", window=3, show_diff = TRUE)
``` 

</br></br>


### Sliding Window of Multivariate Scores
Next we will calculating anomaly score for each lamb, using the "mv" desult scores set. Two scores are calculated, _T2_MCD75_ which seems to work well, and the _EigenvectorCentrality_ method, which does not separate anomalies well for this dataset.

```{r ex4d, fig.height=5}
window<-7
data.last.7D <- data_diff[(dim(data_diff)[2]-window+1):dim(data_diff)[2],]

MultiNav(data_diff_per,type = "snapshot scores", 
         scores = "mv", window=7, show_diff = TRUE)
``` 
</br></br>



