# Charts

MultiNav provides several basic charts that are used as building blocks for the interactive displays. These charts can also be used as stand-alone. <br/>

First let's create some data:

```{r message=FALSE, error=FALSE, fig.height=2, fig.height=3}
library(MultiNav)
library(MASS)
set.seed(1234)
n <- 300
Sigma <- matrix(c(1.0, .80, .50, .20,
                  .80, 1.0, .05, .05,
                  .50, .05, 1.0, .05,
                  .20, .05, .05, 1.0), ncol = 4)
data <- data.frame(mvrnorm(n, Sigma, mu = c(100,100,100,100), empirical = TRUE))
data<- cbind(seq_id=row.names(data), data)

```




## Line Chart

Simple line chart.<br/>

<p class="well" style=" background-color: #F1F1F1; padding: 10px; border: 1px solid #cccccc;">MultiNav(data,type = "line", x= NULL, y= NULL)</p>


**Arguments:** <br/>
data	 - 
Data set in data.table, data.frame or matrix frormat to use for plot.<br/> 
x	 - 
variable name of x values.<br/> 
y	 - 
variable name for y values.<br/> 


**Example:** <br/>


```{r message=FALSE, error=FALSE, fig.height=2, fig.height=3}
MultiNav(data,type = "line","seq_id", "X1")

```

## Scatter Plot
Simple scatter plot chart.

<p class="well" style=" background-color: #F1F1F1; padding: 10px; border: 1px solid #cccccc;">MultiNav(data,type = "scatter", x= NULL, y= NULL)</p>


**Arguments:** <br/>
data	 - 
Data set in data.table, data.frame or matrix frormat to use for plot.<br/> 
x	 - 
variable name of x values.<br/> 
y	 - 
variable name for y values.<br/> 

**Example:** <br/>

```{r message=FALSE, error=FALSE, fig.height=2, fig.height=3}
MultiNav(data,"scatter","X1", "X2")
``` 


## Network Chart
Network chart, with force layout. 

<p class="well" style=" background-color: #F1F1F1; padding: 10px; border: 1px solid #cccccc;">MultiNav(data,type = "network")</p>


**Arguments:** <br/>
data	 - 
Dataset in data.table, data.frame or matrix frormat. The dataset should have columns named: "from", "to", "weight". The weight should be scaled to the numeric range 0-300.<br/> 


**Example:** <br/>		

```{r message=FALSE, error=FALSE, fig.height=3, fig.height=3}
library(igraph)
data <- DendrometerSensors 
cor<-cor(data,use="pairwise.complete.obs")


g  <- graph.adjacency((1-cor)*150,weighted=TRUE, diag=FALSE,mode= "upper")
network <- get.data.frame(g)

head(network)

MultiNav(network,type = "network")


``` 



## Functional BoxPlot
The functional boxplot provides a visual tool for understanding
the behavior of majority of the data over time. In our functional boxplot, the inner envelope, shows the range between the 25th percentile of the readings
over time, and the 75th percentile. The outer envelope is drawn
with a brighter color, and shows the range from the 2nd percentile to
the 98th percentile of the readings over time (containing 96% of the
data). <br/>

<p class="well" style=" background-color: #F1F1F1; padding: 10px; border: 1px solid #cccccc;">MultiNav(data, type = "functional_box",
     raw_data = NULL, line_id = NULL)</p>


**Arguments:** <br/>
data	 - 
Dataset in data.table, data.frame or matrix frormat. The dataset should have columns named: seq_id", "min","quantile_a", "quantile_b","median","quantile_c","quantile_d", "max"

**Optional arguments:** <br/>
raw_data  - Dataset containing the raw data used for calculating the functional boxplot. <br/>
line_id - The column name of the variable to be drawn as a line chart on top of the functional boxplot.

**Example:** <br/>


```{r message=FALSE, error=FALSE, fig.height=2, fig.height=3}

data <- DendrometerSensors
data<-cbind(seq_id=as.numeric(row.names(data)), data)

uni_matrix<-Calc_uni_matrix(as.data.frame(data))
quantiles_matrix<-Calc_quantiles_matrix(as.data.frame(data))

```

```{r message=FALSE, error=FALSE, fig.height=2, fig.height=3}

MultiNav(quantiles_matrix,type = "functional_box")

``` 

<br/>
**Example:** <br/>
```{r message=FALSE, error=FALSE, fig.height=2, fig.height=3}

MultiNav(quantiles_matrix,type = "functional_box_with_line", 
         raw_data = data, line_id="185")

``` 



