# Linked Charts

MultiNav also provides several options for linked charts. Two inputs are required: Multivariate data (raw data) and summary data and. Each id in the summary data will be connected to the mattching variable (with the same id) in the raw table. 

First, let's create some data:

```{r message=FALSE, error=FALSE}
library(MultiNav)
data <- DendrometerSensors
uni_matrix<-Calc_uni_matrix(data)
quantiles_matrix<-Calc_quantiles_matrix(data)

data<-cbind(seq_id=as.numeric(row.names(data)), data)

```

## Scatter Plot With Linked Line Chart

Creates a scatter plot With linked line chart. With an option to add functional boxplot as a background for the line chart.

<p class="well" style=" background-color: #F1F1F1; padding: 10px; border: 1px solid #cccccc;">MultiNav(data, type = "scatter_and_linked_line",
     x= NULL, y= NULL, raw_data = NULL, link_id = NULL, q_data = NULL)</p>


**Arguments:** <br/>
data	 - 
Data set in data.table, data.frame or matrix frormat which contains summary scores for each variable. Includes additional column named: "seq_id" to identify the variable names.<br/>
x	 - 
variable name for the x values of the scatter plot.<br/> 
y	 - 
variable name for the y values of the scatter plot.<br/> 
raw_data - Data set in data.frame frormat containing the raw data of connected variables. <br/> 
link_id - Serves as an identifier for the linked charts (allowing several linked charts on the same markdown page).

**Example:** <br/>
```{r message=FALSE, error=FALSE, fig.height=4}
MultiNav(uni_matrix,type = "scatter_and_linked_line",
         "min", "max", raw_data = data, link_id="chart_a")

``` 

</br></br>
**Optional Arguments:** <br/>
q_data - quantiles dataset, which can be created with aid of the data pre-processing function 'Calc_quantiles_matrix()' 


**Example:** <br/>
Same example as before, with a functional boxplot added in the background of the line chart.

```{r message=FALSE, error=FALSE, fig.height=4}
MultiNav(uni_matrix,type = "scatter_and_linked_line",
         "min", "max", raw_data = data, link_id="chart_b",
         q_data = quantiles_matrix)
``` 

</br></br>


## Network Graph With Linked Line Charts

Creates a network graph with force layout and a linked line chart. With an option to add functional boxplot as a background for the line chart.

<p class="well" style=" background-color: #F1F1F1; padding: 10px; border: 1px solid #cccccc;">MultiNav(data, type = "network_and_linked_line",
     x= NULL, y= NULL, raw_data = NULL, link_id = NULL, q_data = NULL)</p>


**Arguments:** <br/>
data	 - 
Dataset with expected columns named: "from", "to", "weight". The weight should be scaled to the range 0-300.<br/>
x	 - 
variable name for the x values of the scatter plot.<br/> 
y	 - 
variable name for the y values of the scatter plot.<br/> 
raw_data - Dataset containing the raw data of the connected variables. <br/>
link_id - Serves as an id for the linked charts (allowing several linked charts on the same markdown page).

**Optional Arguments:** <br/>
q_data - Data set in data.frame frormat of quantiles over time. Data set can be created with aid of the data pre-processing function 'Calc_quantiles_matrix()' 

**Example:** <br/>

```{r message=FALSE, error=FALSE, fig.height=4}
library(igraph)

exclude.vars <- match(c("seq_id"), names(data)) 
cor<-cor(data[,-exclude.vars ],use="pairwise.complete.obs")

g  <- graph.adjacency((1-cor)*150,weighted=TRUE, diag=FALSE,mode= "upper")
network <- get.data.frame(g)
head(network)
MultiNav(network,type = "network_and_linked_line",
          raw_data = data, link_id="chart_c",
         q_data = quantiles_matrix)

```

## Multiple Linked Charts

There is an option to link multiple charts to the same linked line chart. The multiple link is set by setting the same 'link_id' paramter in multiple charts. This option can be effectvie when knitting rmd files to create dashboard page with many linked charts.


```{r message=FALSE, error=FALSE, fig.height=4}
MultiNav(uni_matrix,type = "scatter_and_linked_line",
         "min", "max", raw_data = data, link_id="chart_D",q_data = quantiles_matrix)

``` 

</br></br>

```{r message=FALSE, error=FALSE, fig.height=2}
MultiNav(uni_matrix,type = "scatter","quantile_25", "quantile_75" , link_id="chart_D")

``` 

